---
title: "Lab 02 - Plastic waste"
author: Lindsey Wilson
date: 1/19/23
output: github_document
---

## Load packages and data

```{r load-packages, message=FALSE}
library(tidyverse) 
```

```{r load-data, message=FALSE}
plastic_waste <- read.csv("data/plastic-waste.csv")
```

## Exercises

### Exercise 1

Here we have the plastic waste per capita separated by continent

```{r plastic-waste-continent}
ggplot(
  data = plastic_waste,
  aes(x = plastic_waste_per_cap),
  color = continent) +
  geom_histogram(binwidth = 0.2) +
  facet_wrap(~ continent)
```

### Exercise 2

#### 2.1

If we include all of the countries, the plot looks like this:

```{r plastic-waste-density-all}
ggplot(data = plastic_waste, 
       mapping = aes(x = plastic_waste_per_cap, 
                     color = continent, 
                     fill = continent)) +
  geom_density(alpha = 0.1)
```

But the one outlier point on the far right is still throwing things off.
By running this code, we can identify that data point as coming from Trinidad & Tobago:

```{r outlier-ID}
plastic_waste %>%
  filter(plastic_waste_per_cap > 3.5)
```

The plot might be a little more useful if we filter Trinidad & Tobago out.
We've done that below (and will continue to do that for the remaining plots):

```{r plastic-waste-density-filtered}
plastic_waste_filtered <- plastic_waste %>%
  filter(code != "TTO")

ggplot(data = plastic_waste_filtered, 
       mapping = aes(x = plastic_waste_per_cap, 
                     color = continent, 
                     fill = continent)) +
  geom_density(alpha = 0.1)
```

#### 2.2

`color` and `fill` are defined by mapping aesthetics of the plot because they're meaningful; we want them to be mapped to the `continent` variable.
Defining those two characteristics as properties of the plot makes sure that they always agree, and allows us to play around with the type of visualization we want to use without breaking that binding.

In contrast `alpha` is a property of the geom because we want it to be the same for all of the curves; we don't want it to be based on a variable.
For example, we don't want the curve for North America to be more transparent than the curve for Africa.

### Exercise 3

Lets' try next to represent out data as a set of violin plots.
These differ from box plots because, while a boxplot only shows summary statistics, violin plots show summary stats *and* the density of the data at each point along the y-axis.

Below we have our data represented with violin plots:

```{r plastic-waste-violin}
ggplot(data = plastic_waste_filtered,
       mapping = aes(x = continent,
                     y = plastic_waste_per_cap,
                     color = continent,
                     fill = continent)) + 
  geom_violin()
```
### Exercise 4

Remove this text, and add your answer for Exercise 4 here.



### Exercise 5

Remove this text, and add your answer for Exercise 5 here.

```{r plastic-waste-mismanaged}
# insert code here
```

### Exercise 6

Remove this text, and add your answer for Exercise 6 here.

```{r plastic-waste-mismanaged-continent}
# insert code here
```

### Exercise 7

Remove this text, and add your answer for Exercise 7 here.

```{r plastic-waste-population-total}
# insert code here
```

```{r plastic-waste-population-coastal}
# insert code here
```

### Exercise 8

Remove this text, and add your answer for Exercise 8 here.

```{r recreate-viz}
# insert code here
```


